---
title: "Data Exploration"
author: "Felix Bigler"
date: "11/12/2021"
output:
  html_document:
    toc: yes
    toc_float: no
    toc_depth: 4
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---

````{r, echo= FALSE}
# load library
library(rstudioapi)
library(ggplot2)
library(gridExtra)
library(dplyr)
```

````{r, echo=FALSE}
# set directory read data
setwd(dirname(getActiveDocumentContext()$path))
df <- read.csv("../Data/hour.csv")
`````
# Graphical Data Exploration

## Influnece of Weather upon Bike Rentals

In this chapter we would like to explore how different weather conditions influence the bike rental numbers.

### Exploration of Continuous Variables

We want to explore how different weather situation influences the number of bike rentals. For that we have following 4 continuous weather variables at hand which we are going to examine:

- temp:       Normalized temperature in Celsius. t_min [-8], t_max [+39]
- atemp:      Normalized feels like temperature in Celsius.  t_min [-16], t_max [+50]
- hum:        Normalized humidity. The values are divided with max. of 100 
- windspeed:  Normalized wind speed. The values are divided with min. of 67 

The continuous weather variables are normalized as follows in the dataset:

````{r,echo=FALSE}
  
str(df)
summary(df)
```

$$
x_{normalized} = \frac{x-x_{min}}{x_{max}-x_{min}}
$$
We assume that we could have used the values without normalization. Since the values are already normalized we will use the as provided in the set.

````{r,echo=FALSE}
  
p1 <-  ggplot(data = df,
         mapping = aes(y = log(cnt),
                     x = temp)) + xlab("Normalized Temperature") + ylab("Num. of Bike Rentals") + 
     geom_point(alpha = 0.1,shape = 1) +
     geom_smooth(aes(colour = "gam")) +
     geom_smooth(aes(colour = "lm"), method = "lm") +
     labs(colour = "Method")

p2 <- ggplot(data = df,
         mapping = aes(y = log(cnt),
                     x = atemp)) +  xlab("Normalized Feels Like Temperature") + ylab("") +
     geom_point(alpha = 0.1,shape = 1) +
     geom_smooth(aes(colour = "gam")) +
     geom_smooth(aes(colour = "lm"), method = "lm") +
     labs(colour = "Method")

p3 <- ggplot(data = df,
         mapping = aes(y = log(cnt),
                     x = hum)) + xlab("Normalized Humidity") + ylab("Num. of Bike Rentals") +
     geom_point(alpha = 0.1,shape = 1) +
     geom_smooth(aes(colour = "gam")) +
     geom_smooth(aes(colour = "lm"), method = "lm") +
     labs(colour = "Method")
 
p4 <- ggplot(data = df,
         mapping = aes(y = log(cnt),
                     x = windspeed)) + xlab("Normalized Windspeed") + ylab("") +
     geom_point(alpha = 0.1,shape = 1) +
     geom_smooth(aes(colour = "gam")) +
     geom_smooth(aes(colour = "lm"), method = "lm") +
     labs(colour = "Method")

grid.arrange(p1,p2,p3,p4, nrow=2,top = "Continuous Weather Variables vs. Number of Bike Rentals")

fig.align = 'center'

```


In all 4 scatter plots the linear model lines (green) as well as the gam model lines (red) are plotted. For both temperatur variables the rentals seems to increase until the temperature reaches a certain value.At the beginning of the obseravtions (low temp.) ans well as the end (high temp.) the linear model seems to overestimate the number of rentals.
For the normalized humidity there is a decrease in rentals with increasing humidity. The is not the case between 0 and 25 % humidity where the linear model overestimates the number of rentals. We assume that in very cold and very hot weather the humidity is rather low and so the number of bike rentals because of the influence of temperature.
In conclusion of the above exploration, it can be stated that the behavior cannot be completely represented with a linear model. There are nonlinear aspects which should be considered in the prediction models.

In the next chapter we would like to examine the categorical variables which are available in the dataset.

### Exploration of Categorical Variables

The data set contains 4 different weather categories which are described as follows:

- Category 1:      Clear, Few clouds, Partly cloudy, Partly cloudy
- Category 2:      Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
- Category 3:        Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds 
- Category 4:  Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog 

Since we are considering the response variables count data we log transform the variable with the natural logarithm in all of the following plots.
````{r,echo=FALSE}

ggplot(data = df,
       mapping = aes(y = log(cnt),
                     x = factor(weathersit),fill= weathersit)) + ylab("Num. of Bike Rentals") + xlab("Weather Conditions") +  labs(fill="Weather\nConditions") +
geom_boxplot() 


```

The boxplot above indicates, that in situation 1 and 2 the most bikes are rented whereas there is seems to be a drop visible  for condition 3 and the drop is even more significant for situation 4 where severe weather conditions are present.This rental behavior is as we have expected since few people ride bicycles in bad weather situations.

In the next chapter we are inspecting the seasonal and timly influence upon the rental situation.

## Influnece of Timely and Seasonal aspects upon Bike Rentals

### Exploration of Categorical Variables


#### Seasonal Influence 


In order to investigate the sesonal influences, the monthly (1 to 12) and seasonal data are included in the dataset as shown below::

- Category 1:      Spring
- Category 2:      Summer
- Category 3:      Fall
- Category 4:      Winter

````{r,echo=FALSE}

p1 <- ggplot(data = df,
       mapping = aes(y = log(cnt),
                     x = factor(season),fill= season)) + ylab("Num. of Bike Rentals") + xlab("Season") +  labs(fill="Season") +
geom_boxplot() 


p2 <- ggplot(data = df,
       mapping = aes(y = log(cnt),
                     x = factor(mnth),fill= mnth)) + ylab("Num. of Bike Rentals") + xlab("Month") +  labs(fill="Monthn") +
geom_boxplot()



grid.arrange(p1,p2, nrow=2,top = "Seasonal Aspects vs. Number of Bike Rentals")

fig.align = 'center'
```

The two boxplot above indicate, that the mos bike rentals take place in summer and fall which we have expected.

#### Timely Influence 

In order to examine the temporal influences we have the time specification in hours (1 to 24), the weekdays (0 = Monday to 6 = Sunday), working day (1 = yes, 0 = no) as well as vacation day (1 = yes , 0 = no).


````{r,echo=FALSE}

p1 <- ggplot(data = df,
       mapping = aes(y = log(cnt),
                     x = factor(hr),fill= hr)) + ylab("Num. of Bike Rentals") + xlab("Hour") +  labs(fill="Hour") +
geom_boxplot() 

p2 <- ggplot(data = df,
       mapping = aes(y = log(cnt),
                     x = factor(weekday),fill= weekday)) + ylab("Num. of Bike Rentals") + xlab("Weekday") +  labs(fill="Day") +
geom_boxplot()

p3 <- ggplot(data = df,
       mapping = aes(y = log(cnt),
                     x = factor(workingday),fill= workingday)) + ylab("Num. of Bike Rentals") + xlab("Workingday") +  labs(fill="Day") +
geom_boxplot() 

p4 <- ggplot(data = df,
       mapping = aes(y = log(cnt),
                     x = factor(holiday),fill= holiday)) + ylab("Num. of Bike Rentals") + xlab("Holidays") +  labs(fill="Day") +
geom_boxplot() 

grid.arrange(p1,p2,p3,p4, nrow=2,top = "Timely Aspects vs. Number of Bike Rentals")


```

The above plots indicate, that the hour in which the bike is rented has the most influence of total rentals which seems obvious. On the other hand, whether there is a working day or not, seems not to have a significant influence.


In the next chapter we try to find a linear model which suits best the dateset.
```








