---
title: "Neural Net"
output: html_notebook
author: "Régis Michael Andreoli"
---


1.1 Set up - Load libraries & Data preparation

2.1 Split data into train and test partition

3.1 First Model - Train & compare - with diversed predictors
3.2 First Model - Test best model
3.2 First Model - Finetune best model
3.3 First Model - Test final model

4.1 Second Model - Train & compare - only as time series
4.2 Second Model - Test best model
4.3 Second Model - Finetune best model
4.4 Second Model - Test final model

5.1 Create SARIMA model and compare with best second model


--------------------------------------------------------------------------------
#1.1 Set up - Load libraries & Data preparation

##1.1 Load libraries
```{r, echo= FALSE}
library(tidyverse)
library(neuralnet)
library(caret)
library(rstudioapi)
library(rmarkdown)
library(dplyr)
```

##1.1 Load data
```{r, echo= FALSE}
d <- setwd(dirname(getActiveDocumentContext()$path))
df <- read.csv("../Data/hour.csv")
```

##1.1 Decompose multi factor variable into single factor variable
```{r, echo= FALSE}
df$season_1 <- ifelse(df$season == "1", 1, 0)
df$season_2 <- ifelse(df$season == "2", 1, 0)
df$season_3 <- ifelse(df$season == "3", 1, 0)
df$season_4 <- ifelse(df$season == "4", 1, 0)
df$weathersit_1 <- ifelse(df$weathersit == "1", 1, 0)
df$weathersit_2 <- ifelse(df$weathersit == "2", 1, 0)
df$weathersit_3 <- ifelse(df$weathersit == "3", 1, 0)
df$weathersit_4 <- ifelse(df$weathersit == "4", 1, 0)
df$hr_0 <- ifelse(df$hr == "0", 1, 0)
df$hr_1 <- ifelse(df$hr == "1", 1, 0)
df$hr_2 <- ifelse(df$hr == "2", 1, 0)
df$hr_3 <- ifelse(df$hr == "3", 1, 0)
df$hr_4 <- ifelse(df$hr == "4", 1, 0)
df$hr_5 <- ifelse(df$hr == "5", 1, 0)
df$hr_6 <- ifelse(df$hr == "6", 1, 0)
df$hr_7 <- ifelse(df$hr == "7", 1, 0)
df$hr_8 <- ifelse(df$hr == "8", 1, 0)
df$hr_9 <- ifelse(df$hr == "9", 1, 0)
df$hr_10 <- ifelse(df$hr == "10", 1, 0)
df$hr_11 <- ifelse(df$hr == "11", 1, 0)
df$hr_12 <- ifelse(df$hr == "12", 1, 0)
df$hr_13 <- ifelse(df$hr == "13", 1, 0)
df$hr_14 <- ifelse(df$hr == "14", 1, 0)
df$hr_15 <- ifelse(df$hr == "15", 1, 0)
df$hr_16 <- ifelse(df$hr == "16", 1, 0)
df$hr_17 <- ifelse(df$hr == "17", 1, 0)
df$hr_18 <- ifelse(df$hr == "18", 1, 0)
df$hr_19 <- ifelse(df$hr == "19", 1, 0)
df$hr_20 <- ifelse(df$hr == "20", 1, 0)
df$hr_21 <- ifelse(df$hr == "21", 1, 0)
df$hr_22 <- ifelse(df$hr == "22", 1, 0)
df$hr_23 <- ifelse(df$hr == "23", 1, 0)
df$weekday_1 <- ifelse(df$weekday == "0", 1, 0)
df$weekday_2 <- ifelse(df$weekday == "1", 1, 0)
df$weekday_3 <- ifelse(df$weekday == "2", 1, 0)
df$weekday_4 <- ifelse(df$weekday == "3", 1, 0)
df$weekday_5 <- ifelse(df$weekday == "4", 1, 0)
df$weekday_6 <- ifelse(df$weekday == "5", 1, 0)
df$weekday_7 <- ifelse(df$weekday == "6", 1, 0)
df$mnth_1 <- ifelse(df$mnth == "1", 1, 0)
df$mnth_2 <- ifelse(df$mnth == "2", 1, 0)
df$mnth_3 <- ifelse(df$mnth == "3", 1, 0)
df$mnth_4 <- ifelse(df$mnth == "4", 1, 0)
df$mnth_5 <- ifelse(df$mnth == "5", 1, 0)
df$mnth_6 <- ifelse(df$mnth == "6", 1, 0)
df$mnth_7 <- ifelse(df$mnth == "7", 1, 0)
df$mnth_8 <- ifelse(df$mnth == "8", 1, 0)
df$mnth_9 <- ifelse(df$mnth == "9", 1, 0)
df$mnth_10 <- ifelse(df$mnth == "10", 1, 0)
df$mnth_11 <- ifelse(df$mnth == "11", 1, 0)
df$mnth_12 <- ifelse(df$mnth == "12", 1, 0)
```

##1.1 Normalize dependent var
```{r}
data <- df
data$cnt <- (df$cnt - min(df$cnt)) / (max(df$cnt) - min(df$cnt))
str(data)
```
