---
title: "Neural Net"
output: html_notebook
author: "Régis Michael Andreoli"
---


#1.1 Set up - Load libraries & Data preparation

#2.1 Split data into train and test partition

#3.1 First Model - Train & compare - with diversed predictors
#3.2 First Model - Test best model
#3.2 First Model - Finetune best model
#3.3 First Model - Test final model

#4.1 Second Model - Train & compare - only as time series
#4.2 Second Model - Test best model
#4.3 Second Model - Finetune best model
#4.4 Second Model - Test final model

#5.1 Create SARIMA model and compare with best second model


--------------------------------------------------------------------------------
#1.1 Set up - Load libraries & Data preparation

##1.1 Load libraries
```{r, echo= FALSE}
library(tidyverse)
library(neuralnet)
library(caret)
library(rstudioapi)
library(rmarkdown)
library(dplyr)
```

##1.1 Load data
```{r, echo= FALSE}
d <- setwd(dirname(getActiveDocumentContext()$path))
df <- read.csv("../Data/hour.csv")
```

##1.1 Decompose multi factor variable into single factor variable
```{r, echo= FALSE}
df$season_1 <- ifelse(df$season == "1", 1, 0)
df$season_2 <- ifelse(df$season == "2", 1, 0)
df$season_3 <- ifelse(df$season == "3", 1, 0)
df$season_4 <- ifelse(df$season == "4", 1, 0)
```

##1.1 Adjust variables types
```{r, echo= FALSE}
df$season_1 <- as.factor(df$season_1)
df$season_2 <- as.factor(df$season_2)
df$season_3 <- as.factor(df$season_3)
df$season_4 <- as.factor(df$season_4)
df$season <- as.factor(df$season)
df$yr <- as.factor(df$yr)
df$mnth <- as.factor(df$mnth)
df$hr <- as.factor(df$hr)
df$holiday <- as.factor(df$holiday)
df$weekday <- as.factor(df$weekday)
df$workingday <- as.factor(df$workingday)
df$weathersit <- as.factor(df$weathersit)
```

##1.1 Normalize dependent var
```{r}
df$cnt <- (df$cnt - min(df$cnt)) / (max(df$cnt) - min(df$cnt))
str(df)
```

##1.1 Identifying near-zero variance variable
```{r}
nearZeroVar(df, saveMetrics = TRUE)
```

#2.1 Split data into train and test partition

```{r}
set.seed(123)
indices <- createDataPartition(df$cnt, p=.3, list = F)

train <- df %>% slice(indices)
test <- df %>% slice(-indices)

str(train)
```

#3.1 First Model - Train & compare - with diversed predictors


##3.1 Set up model A
```{r}
set.seed(44)
tuGrid <- expand.grid(.layer1=c(1:4), .layer2=c(0:4), .layer3=c(0:2))

trCtrl <- trainControl(
  method = 'repeatedcv',
  number = 10,
  repeats = 5,
  returnResamp = 'final'
)
```

##3.1 Train model A
```{r}
models <- train(cnt ~ weekday + weathersit, data = train, method = 'neuralnet', metric = 'RMSE',
  linear.output = TRUE,
  preProcess = c('center', 'scale'),
  tuneGrid = tuGrid,
  trControl = trCtrl
  )
```

```{r}
plot(models)
```

```{r}
plot(models_$finalModel)
```

# ANN Model 2 - With Caret - Model to predict amount of rents only due to date variables

##3.1 Set up model B
```{r}
set.seed(44)
tuGrid_2 <- expand.grid(.layer1=c(1:x), .layer2=c(x:x), .layer3=c(0:x))

trCtrl_2 <- trainControl(
  method = 'repeatedcv',
  number = 10,
  repeats = 5,
  returnResamp = 'final'
)
```

##3.1 Train model B
```{r}
models_2 <- train(cnt ~ hr + weekday + mnth, data = train, method = 'neuralnet', metric = 'RMSE',
  linear.output = TRUE,
  preProcess = c('center', 'scale'),
  tuneGrid = tuGrid_2,
  trControl = trCtrl_2
  )
```

```{r}
plot(models_2)
```
```{r}
plot(models_2$finalModel)
```

# Compute prediction with test data. Then denormalized data
```{r}
pred_df <- compute(models_2$finalModel, test$cnt)
pred <- pred_df$net.result * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
pred
```

# Train best Neuron&Layer model with larger dataset

##Split data in train & test partition
```{r}
set.seed(123)
indices_2 <- createDataPartition(df$cnt, p=.8, list = F)

train_2 <- df %>% slice(indices_2)
test_2 <- df %>% slice(-indices_2)

str(train)
```

```{r}
set.seed(42)
final_model = neuralnet(cnt ~ temp + windspeed, train_2, hidden = 3)
```
```{r}
plot(final_model)
```



*** under is crap ***



```{r}
set.seed(44)
tuGrid_3 <- expand.grid(.layer1=c(3), .layer2=c(5), .layer3=c(2))

trCtrl_3 <- trainControl(
  method = 'repeatedcv',
  number = 10,
  repeats = 5,
  returnResamp = 'final'
)
```

```{r}
models_2 <- train(cnt ~ hr + weekday + mnth, data = train,
  method = 'neuralnet', metric = 'RMSE',
  linear.output = TRUE,
  preProcess = c('center', 'scale'),
  tuneGrid = tuGrid_3,
  trControl = trCtrl_3
  )
```

