---
title: "Neural Net"
output: html_notebook
author: "Régis Michael Andreoli"
---


1.1 Set up - Load libraries & Data preparation

2.1 Split data into train and test partition

3.1 First Model - Train & compare - with diversed predictors
3.2 First Model - Test best model
3.2 First Model - Finetune best model
3.3 First Model - Test final model

4.1 Second Model - Train & compare - only as time series
4.2 Second Model - Test best model
4.3 Second Model - Finetune best model
4.4 Second Model - Test final model

5.1 Create SARIMA model and compare with best second model


--------------------------------------------------------------------------------
#1.1 Set up - Load libraries & Data preparation

##1.1 Load libraries
```{r, echo= FALSE}
library(tidyverse)
library(neuralnet)
library(caret)
library(rstudioapi)
library(rmarkdown)
library(dplyr)
```

##1.1 Load data
```{r, echo= FALSE}
d <- setwd(dirname(getActiveDocumentContext()$path))
df <- read.csv("../Data/hour.csv")
```

##1.1 Decompose multi factor variable into single factor variable
```{r, echo= FALSE}
df$season_1 <- ifelse(df$season == "1", 1, 0)
df$season_2 <- ifelse(df$season == "2", 1, 0)
df$season_3 <- ifelse(df$season == "3", 1, 0)
df$season_4 <- ifelse(df$season == "4", 1, 0)
df$weathersit_1 <- ifelse(df$weathersit == "1", 1, 0)
df$weathersit_2 <- ifelse(df$weathersit == "2", 1, 0)
df$weathersit_3 <- ifelse(df$weathersit == "3", 1, 0)
df$weathersit_4 <- ifelse(df$weathersit == "4", 1, 0)
df$hr_0 <- ifelse(df$hr == "0", 1, 0)
df$hr_1 <- ifelse(df$hr == "1", 1, 0)
df$hr_2 <- ifelse(df$hr == "2", 1, 0)
df$hr_3 <- ifelse(df$hr == "3", 1, 0)
df$hr_4 <- ifelse(df$hr == "4", 1, 0)
df$hr_5 <- ifelse(df$hr == "5", 1, 0)
df$hr_6 <- ifelse(df$hr == "6", 1, 0)
df$hr_7 <- ifelse(df$hr == "7", 1, 0)
df$hr_8 <- ifelse(df$hr == "8", 1, 0)
df$hr_9 <- ifelse(df$hr == "9", 1, 0)
df$hr_10 <- ifelse(df$hr == "10", 1, 0)
df$hr_11 <- ifelse(df$hr == "11", 1, 0)
df$hr_12 <- ifelse(df$hr == "12", 1, 0)
df$hr_13 <- ifelse(df$hr == "13", 1, 0)
df$hr_14 <- ifelse(df$hr == "14", 1, 0)
df$hr_15 <- ifelse(df$hr == "15", 1, 0)
df$hr_16 <- ifelse(df$hr == "16", 1, 0)
df$hr_17 <- ifelse(df$hr == "17", 1, 0)
df$hr_18 <- ifelse(df$hr == "18", 1, 0)
df$hr_19 <- ifelse(df$hr == "19", 1, 0)
df$hr_20 <- ifelse(df$hr == "20", 1, 0)
df$hr_21 <- ifelse(df$hr == "21", 1, 0)
df$hr_22 <- ifelse(df$hr == "22", 1, 0)
df$hr_23 <- ifelse(df$hr == "23", 1, 0)
df$weekday_1 <- ifelse(df$weekday == "0", 1, 0)
df$weekday_2 <- ifelse(df$weekday == "1", 1, 0)
df$weekday_3 <- ifelse(df$weekday == "2", 1, 0)
df$weekday_4 <- ifelse(df$weekday == "3", 1, 0)
df$weekday_5 <- ifelse(df$weekday == "4", 1, 0)
df$weekday_6 <- ifelse(df$weekday == "5", 1, 0)
df$weekday_7 <- ifelse(df$weekday == "6", 1, 0)
df$mnth_1 <- ifelse(df$mnth == "1", 1, 0)
df$mnth_2 <- ifelse(df$mnth == "2", 1, 0)
df$mnth_3 <- ifelse(df$mnth == "3", 1, 0)
df$mnth_4 <- ifelse(df$mnth == "4", 1, 0)
df$mnth_5 <- ifelse(df$mnth == "5", 1, 0)
df$mnth_6 <- ifelse(df$mnth == "6", 1, 0)
df$mnth_7 <- ifelse(df$mnth == "7", 1, 0)
df$mnth_8 <- ifelse(df$mnth == "8", 1, 0)
df$mnth_9 <- ifelse(df$mnth == "9", 1, 0)
df$mnth_10 <- ifelse(df$mnth == "10", 1, 0)
df$mnth_11 <- ifelse(df$mnth == "11", 1, 0)
df$mnth_12 <- ifelse(df$mnth == "12", 1, 0)
```

##1.1 Normalize dependent var
```{r}
data <- df
data$cnt <- (df$cnt - min(df$cnt)) / (max(df$cnt) - min(df$cnt))
str(data)
```

##1.1 Identifying near-zero variance variable
```{r}
nearZeroVar(data, saveMetrics = TRUE)
```

#3.1 Model 1

##3.1 Split data into train and test partition
```{r}
set.seed(123)
indices_1 <- createDataPartition(data$cnt, p=.65, list = F)

train_1 <- data %>% slice(indices_1)
test_1 <- data %>% slice(-indices_1)

str(train_1)
```


##3.1 Model 1 - Set up
```{r}
set.seed(44)
tuGrid_1 <- expand.grid(.layer1=c(1,2,4:8), .layer2=c(0,2,3,4), .layer3=c(0,2))

trCtrl_1 <- trainControl(
  method = 'repeatedcv',
  number = 5,
  repeats = 1,
  returnResamp = 'final'
)
```

##3.1 Model 1 - Train and Compare
```{r}#
models_1 <- train(cnt ~ hum + temp + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23, data = train,
  method = 'neuralnet',
  metric = 'RMSE',
  linear.output = TRUE,
  threshold = 0.1,
  lifesign.step = 1000,
  lifesign = "full",
  preProcess = c('center', 'scale'),
  tuneGrid = tuGrid_1,
  trControl = trCtrl_1
  )
```

##3.1 Model 1 - Save and Load
```{r}#
saveRDS(models_1, "neural_nets_models_1.rds")
```

```{r}
models_1 <-readRDS("neural_nets_models_1.rds")
```

```{r}
plot(models_1)
```

##3.1 Model 1 - Compute Prediction
```{r}
pred_1 <- compute(models_1$finalModel, test_1 %>% select(-cnt))
pred_1 <- pred_1$net.result * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
control_1 <- test_1$cnt * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
head(pred_1)
head(control_1)
```

##3.1 Model 1 - Root Mean Square
```{r}
sqrt(mean((control_1 - pred_1)^2))
```

##3.1 Model 1 - Plot Model Performance
```{r}
plot(control_1, pred_1, col='orange', cex = .3, pch=20, ylab = "predicted rating NN", xlab = "real rating")
abline(0,1)
```

#3.2 Model 2

##3.2 Split Data into Train and Test Partition
```{r}
set.seed(42)
indices_2 <- createDataPartition(data$cnt, p=.8, list = F)

train_2 <- data %>% slice(indices_2)
test_2 <- data %>% slice(-indices_2)
```

##3.2 Model 2 - Train
```{r}#
set.seed(42)
model_2 = neuralnet(cnt ~ hum + temp + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23, data = train_2, hidden = 4,linear.output = TRUE, threshold = 0.01, stepmax = 500000, lifesign.step = 1000, lifesign = "full")
```

##3.2 Model 2 - Save and Load
```{r}#
saveRDS(model_2, "neural_nets_model_2.rds")
```

```{r}
model_2 <-readRDS("neural_nets_model_2.rds")
```

##3.2 Model 2 - Compute Prediction
```{r}
pred_2 <- compute(model_2, test_2 %>% select(-cnt))
pred_2 <- pred_2$net.result * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
control_2 <- test_2$cnt * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
head(pred_2)
head(control_2)
```

##3.2 Model 2 - Root Mean Square
```{r}
sqrt(mean((control_2 - pred_2)^2))
```

##3.2 Model 2 - Plot Model Performance
```{r}
plot(control_2, pred_2, col='orange', cex = .3, pch=20, ylab = "predicted rating NN", xlab = "real rating")
abline(0,1)
```

#3.3 Model 3

##3.3 Model 3 - Train
```{r}#
set.seed(42)
model_3 = neuralnet(cnt ~ hum + temp + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23, data = train_2, hidden = c(7,2,2),linear.output = TRUE, threshold = 0.01, stepmax = 500000, lifesign.step = 1000, lifesign = "full")
```

##3.3 Model 3 - Save and Load
```{r}#
saveRDS(model_3, "neural_nets_model_3.rds")
```

```{r}
model_3 <-readRDS("neural_nets_model_3.rds")
```

##3.3 Model 3 - Compute Prediction
```{r}
pred_3 <- compute(model_3, test_2 %>% select(-cnt))
pred_3 <- pred_3$net.result * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
control_3 <- test_2$cnt * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
head(pred_3)
head(control_3)
```

##3.3 Model 3 - Root Mean Square
```{r}
sqrt(mean((control_3 - pred_3)^2))
```

##3.3 Model 3 - Plot Model Performance
```{r}
plot(control_3, pred_3, col='orange', cex = .3, pch=20, ylab = "predicted rating NN", xlab = "real rating")
abline(0,1)
```

#3.4 Model 4 - all variable - low model complexity

##3.4 Model 4 - set up
```{r}
set.seed(41)
model_4 = neuralnet(cnt ~ hum + temp + windspeed + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23 + weekday_1 + weekday_2 + weekday_3 + weekday_4 + weekday_5 + weekday_6 + weekday_7 + mnth_1 + mnth_2 + mnth_3 + mnth_4 + mnth_5 + mnth_6 + mnth_7 + mnth_8 + mnth_9 + mnth_10 + mnth_11 + mnth_12, data = train_2, hidden = c(4),linear.output = TRUE, threshold = 0.05, stepmax = 600000, lifesign.step = 1000, lifesign = "full")
```

##3.4 Model 4 - Save and Load
```{r}#
saveRDS(model_4, "neural_nets_model_4.rds")
```

```{r}
model_4 <-readRDS("neural_nets_model_4.rds")
```

predict.nn train_2?
##3.4 Model 4 - Compute Prediction
```{r}
pred_4 <- compute(model_4, test_2 %>% select(-cnt))
pred_4 <- pred_4$net.result * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
control_4 <- test_2$cnt * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
head(pred_4)
head(control_4)
```

##3.4 Model 4 - Root Mean Square
```{r}
sqrt(mean((control_4 - pred_4)^2))
```

##3.4 Model 4 - Plot Model Performance
```{r}
plot(control_4, pred_4, col='orange', cex = .3, pch=20, ylab = "predicted rating NN", xlab = "real rating")
abline(0,1)
```

confidence.interval(x, alpha = 0.05)

confidence.interval(model)

gwplot(x, rep = NULL, max = NULL, min = NULL, file = NULL,
selected.covariate = 1, selected.response = 1, highlight = FALSE,
type = "p", col = "black", ...)

gwplot(net.infert, selected.covariate="parity")
gwplot(net.infert, selected.covariate="induced")
gwplot(net.infert, selected.covariate="spontaneous")

#3.5 Models 5

##3.5 Models 5 - Set up
```{r}
set.seed(44)
tuGrid_2 <- expand.grid(.layer1=c(4:8), .layer2=c(0,2,3,4), .layer3=c(0,2))

trCtrl_2 <- trainControl(
  method = 'repeatedcv',
  number = 5,
  repeats = 1,
  returnResamp = 'final',
)
```

##3.5 Models 5 - Train and Compare
```{r}
models_5 <- train(cnt ~ hum + temp + windspeed + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23 + weekday_1 + weekday_2 + weekday_3 + weekday_4 + weekday_5 + weekday_6 + weekday_7 + mnth_1 + mnth_2 + mnth_3 + mnth_4 + mnth_5 + mnth_6 + mnth_7 + mnth_8 + mnth_9 + mnth_10 + mnth_11 + mnth_12, data = train_2,
  method = 'neuralnet',
  metric = 'RMSE',
  linear.output = TRUE,
  threshold = 0.1,
  stepmax = 250000,
  lifesign.step = 1000,
  lifesign = "full",
  preProcess = c('center', 'scale'),
  tuneGrid = tuGrid_2,
  trControl = trCtrl_2
  )
```

```{r}
saveRDS(models_5, "neural_nets_models_5.rds")
```


```{r}
plot(models_5)
```

##3.5 Model 5 - Compute Prediction
```{r}
pred_5 <- compute(models_5$finalModel, test_2 %>% select(-cnt))
pred_5 <- pred_5$net.result * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
control_5 <- test_2$cnt * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
head(pred_5)
head(control_5)
```

##3.5 Model 5 - Root Mean Square
```{r}
sqrt(mean((control_5 - pred_5)^2))
```

##3.5 Model 5 - Plot Model Performance
```{r}
plot(control_5, pred_5, col='orange', cex = .3, pch=20, ylab = "predicted rating NN", xlab = "real rating")
abline(0,1)
```
#Model 6
##3.6 Model 6 - set up
```{r}
set.seed(41)
model_6 = neuralnet(cnt ~ hum + temp + windspeed + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23 + weekday_1 + weekday_2 + weekday_3 + weekday_4 + weekday_5 + weekday_6 + weekday_7 + mnth_1 + mnth_2 + mnth_3 + mnth_4 + mnth_5 + mnth_6 + mnth_7 + mnth_8 + mnth_9 + mnth_10 + mnth_11 + mnth_12, data = train_2, hidden = c(7,3),linear.output = TRUE, threshold = 0.01, stepmax = 750000, lifesign.step = 500, lifesign = "full")
```

```{r}#
saveRDS(model_6, "neural_nets_model_6.rds")
```

##3.6 Model 6 - Compute Prediction
```{r}
pred_6 <- compute(model_6, test_2 %>% select(-cnt))
pred_6 <- pred_6$net.result * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
control_6 <- test_2$cnt * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
head(pred_6)
head(control_6)
```

##3.6 Model 6 - Root Mean Square
```{r}
sqrt(mean((control_6 - pred_6)^2))
```

##3.6 Model 6 - Plot Model Performance
```{r}
plot(control_6, pred_6, col='orange', cex = .3, pch=20, ylab = "predicted rating NN", xlab = "real rating")
abline(0,1)
```

#Model 7
##3.7 Model 7 - set up
```{r}
set.seed(41)
model_7 = neuralnet(cnt ~ hum + temp + windspeed + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23 + weekday_1 + weekday_2 + weekday_3 + weekday_4 + weekday_5 + weekday_6 + weekday_7 + mnth_1 + mnth_2 + mnth_3 + mnth_4 + mnth_5 + mnth_6 + mnth_7 + mnth_8 + mnth_9 + mnth_10 + mnth_11 + mnth_12, data = train_2, hidden = c(8,3),linear.output = TRUE, threshold = 0.01, stepmax = 750000, lifesign.step = 500, lifesign = "full")
```

```{r}#
saveRDS(model_7, "neural_nets_model_7.rds")
```

##3.7 Model 7 - Compute Prediction
```{r}
pred_7 <- compute(model_7, test_2 %>% select(-cnt))
pred_7 <- pred_7$net.result * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
control_7 <- test_2$cnt * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
head(pred_7)
head(control_7)
```

##3.7 Model 7 - Root Mean Square
```{r}
sqrt(mean((control_7 - pred_7)^2))
```

##3.7 Model 7 - Plot Model Performance
```{r}
plot(control_7, pred_7, col='orange', cex = .3, pch=20, ylab = "predicted rating NN", xlab = "real rating")
abline(0,1)
```

#3.8 Model 4.2 - all variable - low model complexity

##3.8 Model 4.2 - set up
```{r}
set.seed(41)
model_8 = neuralnet(cnt ~ hum + temp + windspeed + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23 + weekday_1 + weekday_2 + weekday_3 + weekday_4 + weekday_5 + weekday_6 + weekday_7 + mnth_1 + mnth_2 + mnth_3 + mnth_4 + mnth_5 + mnth_6 + mnth_7 + mnth_8 + mnth_9 + mnth_10 + mnth_11 + mnth_12, data = train_2, hidden = c(4),linear.output = TRUE, threshold = 0.005, stepmax = 750000, lifesign.step = 500, lifesign = "full")
```

#3.9 Model 9 - all variable plus year

##3.9x Model 9x - set up
```{r}
set.seed(41)
model_9 = neuralnet(cnt ~ yr + hum + temp + windspeed + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23 + weekday_1 + weekday_2 + weekday_3 + weekday_4 + weekday_5 + weekday_6 + weekday_7 + mnth_1 + mnth_2 + mnth_3 + mnth_4 + mnth_5 + mnth_6 + mnth_7 + mnth_8 + mnth_9 + mnth_10 + mnth_11 + mnth_12, data = train_2, hidden = c(7,3),linear.output = TRUE, threshold = 0.01, stepmax = 750000, lifesign.step = 500, lifesign = "full")
```

```{r}°
saveRDS(model_9, "neural_nets_model_9.rds")
```

##3.9 Model 9 - Compute Prediction
```{r}
pred_9 <- compute(model_9, test_2 %>% select(-cnt))
pred_9 <- pred_9$net.result * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
control_9 <- test_2$cnt * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
head(pred_9)
head(control_9)
```

##3.9 Model 9 - Root Mean Square
```{r}
sqrt(mean((control_9 - pred_9)^2))
```

##3.9 Model 9 - Plot Model Performance
```{r}
plot(control_9, pred_9, col='orange', cex = .3, pch=20, ylab = "predicted rating NN", xlab = "real rating")
abline(0,1)
```
```{r}
set.seed(44)
tuGrid_2 <- expand.grid(.layer1=c(4:8), .layer2=c(0,2,3,4), .layer3=c(0,2))

trCtrl_2 <- trainControl(
  method = 'repeatedcv',
  number = 5,
  repeats = 1,
  returnResamp = 'final',
)
```

##3.10 Models 10 - Set up
```{r}
set.seed(41)
tuGrid_3 <- expand.grid(.layer1=c(6:12), .layer2=c(2:5), .layer3=c(0,2,3))

trCtrl_3 <- trainControl(
  method = 'repeatedcv',
  number = 3,
  repeats = 1,
  returnResamp = 'final',
)
```

##3.10 Models 10 - Train and Compare
```{r}
models_10 <- train(cnt ~ workingday + holiday + yr + hum + temp + windspeed + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23 + weekday_1 + weekday_2 + weekday_3 + weekday_4 + weekday_5 + weekday_6 + weekday_7 + mnth_1 + mnth_2 + mnth_3 + mnth_4 + mnth_5 + mnth_6 + mnth_7 + mnth_8 + mnth_9 + mnth_10 + mnth_11 + mnth_12, data = train_2,
  method = 'neuralnet',
  metric = 'RMSE',
  linear.output = TRUE,
  threshold = 0.05,
  stepmax = 450000,
  lifesign.step = 1000,
  lifesign = "full",
  preProcess = c('center', 'scale'),
  tuneGrid = tuGrid_3,
  trControl = trCtrl_3
  )
```

##3.10 Model 10 - Save and Load
```{r}#
saveRDS(models_10, "neural_nets_models_10.rds")
```

```{r}
plot(models_10)
```


#3.11 Model 11

##3.11 Model 11 - Train model
```{r}#
set.seed(41)
model_11 = neuralnet(cnt ~ workingday + holiday + yr + hum + temp + windspeed + weathersit_1 + weathersit_2 + weathersit_3 + hr_1 + hr_2 + hr_3 + hr_4 + hr_5 + hr_6 + hr_7 + hr_8 + hr_9 + hr_10 + hr_11 + hr_12 + hr_13 + hr_14 + hr_15 + hr_16 + hr_17 + hr_18 + hr_19 + hr_20 + hr_21 + hr_22 + hr_23 + weekday_1 + weekday_2 + weekday_3 + weekday_4 + weekday_5 + weekday_6 + weekday_7 + mnth_1 + mnth_2 + mnth_3 + mnth_4 + mnth_5 + mnth_6 + mnth_7 + mnth_8 + mnth_9 + mnth_10 + mnth_11 + mnth_12, data = train_2, hidden = c(7,3),linear.output = TRUE, threshold = 0.01, stepmax = 750000, lifesign.step = 500, lifesign = "full")
```

```{r}
models_11 <-readRDS("neural_nets_model_11.rds")
```

##3.11 Model 11 - Compute Prediction
```{r}
pred_11 <- compute(model_9x, test_2 %>% select(-cnt))
pred_11 <- pred_11$net.result * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
control_11 <- test_2$cnt * (max(df$cnt) - min(df$cnt)) + min(df$cnt)
head(pred_11)
head(control_11)
```

##3.11 Model 11 - Root Mean Square
```{r}
sqrt(mean((control_11 - pred_11)^2))
```

##3.11 Model 11 - Plot Model Performance
```{r}
plot(control_11, pred_11, col='orange', cex = .3, pch=20, ylab = "predicted rating NN", xlab = "real rating")
abline(0,1)
```

data_ts <- ts(df$cnt)
x <- stl(data_ts, s.window = "periodic")
plot(x)


arima_model <- auto.arima(data_ts,ic="aic")
print(arima_model)

pred <- predict(model, n.ahead = 30)
plot(pred$pred)
lines(real_d, col = "orange")
